### 2022-12-14 manuscript.Rmd
Preprocessing code added

---
title: "Uncanny Valley"
author: "Daniel Baker"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  pdf_document:
    toc: no
  html_document: default
bibliography: references.bib
csl: elife.csl
---

```{r setup, include=FALSE}

processdata <- 3  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data (requires XMB of storage)
# 2 - perform multivariate pattern analysis on individual participants (requires 25GB of storage, and takes X hours)
# 3 - download all raw data and do preprocessing (requires ~60GB of storage)

niterations <- 100  # number of iterations for MVPA

# check which packages are installed, install the missing ones, and activate
packagelist <- c('knitr','e1071','signal','osfr') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

rawdir <- 'local/raw/'
rawdir <- '/Volumes/USB128/UncannyValley/'
processeddir <- 'local/processed/'
processeddir <- '/Volumes/USB128/UncannyValley/'

knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

# Introduction

# Materials & Methods

## Participants

## Apparatus & stimuli

## Procedure

### Experiment 1: robots

### Experiment 2: hyper-realistic masks

## Data analysis

## Data and code availability

Raw data, processed data, and all analysis scripts are available through the project repository at: https://.

# Results

## Experiment 1

```{r include=FALSE, results='hide'}

# (download and) analyse individual participant data for experiment 1

if (processdata > 2){

  legaltriggers <- 101:103
  sublist <- dir(paste0(rawdir,'Robots'),full.names=FALSE)
  subdirs <- dir(paste0(rawdir,'Robots'),full.names=TRUE)

  for (s in 1:length(sublist)){

   if (!file.exists(paste0(processeddir,'RobotsP/R',s,'_processed.RData'))){
    datafiles <- dir(subdirs[s],pattern='*.csv.gz',full.names=TRUE)

    trialcounter <- legaltriggers*0
    alltrials <- array(0,dim=c(3,180,64,1200))
    for (block in 1:length(datafiles)){

      EEGdata <- read.csv(datafiles[block])
      electrodes <- colnames(EEGdata)[3:68]

      # bandpass filter data
      filt <- butter(10,1/16.666,type='low')
      for (ch in 1:66){
        temp <- EEGdata[,ch+2]
        ftemp <- signal::filter(filt,temp)
        EEGdata[,ch+2] <- ftemp
      }

    triggertimes <- NULL
    counter <- 0
    lasttrigger <- -10000
    for (n in 1:nrow(EEGdata)){
      if (EEGdata$Trigger[n] %in% legaltriggers){
        if (n>(lasttrigger+100)){
        counter <- counter + 1
        triggertimes[counter] <- n
        lasttrigger <- n
      }}
    }

    for (trial in 1:counter){
      thiscond <- which(legaltriggers==EEGdata$Trigger[triggertimes[trial]])
      trialcounter[thiscond] <- trialcounter[thiscond] + 1

      for (ch in 1:64){
       temp <- EEGdata[triggertimes[trial]+(-200:999),ch+2]
       temp <- temp - mean(temp[1:200])
       alltrials[thiscond,trialcounter[thiscond],ch,] <- temp
      }
    }

    }

    save(file=paste0(processeddir,'RobotsP/R',s,'_processed.RData'),list=c('trialcounter','alltrials'))
   }
  }

}

```

```{r include=FALSE, results='hide'}

# do MVPA for experiment 1

if (processdata > 1){

}

```

```{r include=FALSE, results='hide'}

# create figures for experiment 1

if (processdata > 0){

}

```

## Experiment 2

```{r include=FALSE, results='hide'}

# (download and) analyse individual participant data for experiment 2

if (processdata > 2){

  # 100: asian, 200: western
  # 11: halloween, 12: silicone, 23: real
  legaltriggers <- c(111,112,123,211,212,223)
  sublist <- dir(paste0(rawdir,'Masks'),full.names=FALSE)
  subdirs <- dir(paste0(rawdir,'Masks'),full.names=TRUE)

  for (s in 1:length(sublist)){

    if (!file.exists(paste0(processeddir,'MasksP/M',s,'_processed.RData'))){
    datafiles <- dir(subdirs[s],pattern='*.csv.gz',full.names=TRUE)

    trialcounter <- matrix(0,nrow=3,ncol=6)
    alltrials <- array(0,dim=c(3,6,76,64,1200))
    for (block in 1:3){

      EEGdata <- read.csv(datafiles[block])
      electrodes <- colnames(EEGdata)[3:68]

      # bandpass filter data
      filt <- butter(10,1/16.666,type='low')
      for (ch in 1:66){
        temp <- EEGdata[,ch+2]
        ftemp <- signal::filter(filt,temp)
        EEGdata[,ch+2] <- ftemp
      }

    triggertimes <- NULL
    counter <- 0
    lasttrigger <- -10000
    for (n in 1:nrow(EEGdata)){
      if (EEGdata$Trigger[n] %in% legaltriggers){
        if (n>(lasttrigger+100)){
        counter <- counter + 1
        triggertimes[counter] <- n
        lasttrigger <- n
      }}
    }

    for (trial in 1:counter){
      thiscond <- which(legaltriggers==EEGdata$Trigger[triggertimes[trial]])
      trialcounter[block,thiscond] <- trialcounter[block,thiscond] + 1

      for (ch in 1:64){
       temp <- EEGdata[triggertimes[trial]+(-200:999),ch+2]
       temp <- temp - mean(temp[1:200])
       alltrials[block,thiscond,trialcounter[block,thiscond],ch,] <- temp
      }
    }

    }

    save(file=paste0(processeddir,'MasksP/M',s,'_processed.RData'),list=c('trialcounter','alltrials'))
  }
  }
}

```

```{r include=FALSE, results='hide'}

# do MVPA for experiment 2

if (processdata > 1){

}

```

```{r include=FALSE, results='hide'}

# create figures for experiment 2

if (processdata > 0){

}

```

```{r pupildata, fig.cap="Caption.", fig.align="center", echo=FALSE}

# knitr::include_graphics(paste0('Figures/pupildata.pdf'))

```
# Discussion

# Conclusions

# Acknowledgements

# References